---
title: "R02_2 Pivot analysis on tp theft with dplyr"
author: "Jilung Hsieh"
date: "2019/9/2"
output:
  html_document:
    theme: cerulean
    highlight: zenburn
    toc: yes
    toc_float:
      collapsed: no
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# loading packages
```{r}
library(tidyverse)
options(stringsAsFactors = F)
```

```{r}
library(httr)
options(stringsAsFactors = F)
url <- "https://data.taipei/api/getDatasetInfo/downloadResource?id=68785231-d6c5-47a1-b001-77eec70bec02&rid=93d9bc2d-af08-4db7-a56b-9f0a49226fa3"
df <- read_csv(url, locale = locale(encoding = "Big5"))
```

# Cleaning data I
- Renaming variables by `select()`

    select(id = 編號, 
           cat = 案類, 
           date = `發生(現)日期`, 
           time = `發生時段`, 
           location = `發生(現)地點`) %>%

- Generating variable year
- Generating variable month
- Retrieving area


```{r}
df1 <- select(df, id = "編號", cat = "案類", date = "發生(現)日期", time = "發生時段", location = "發生(現)地點")

df2 <- mutate(df1, year = date%/%10000)
df3 <- mutate(df2, month = date%%10000%/%100)
df4 <- mutate(df3, area = str_sub(location, 4, 6))
df5 <- mutate(df4, county = str_sub(location, 1, 3))

```

```{r}
df_new <- df %>%
  select(id = "編號", cat = "案類", date = "發生(現)日期", time = "發生時段", location = "發生(現)地點") %>%
  mutate(year = date%/%10000) %>%
  mutate(month = date%%10000%/%100) %>%
  mutate(area = str_sub(location, 4, 6)) %>%
  mutate(county = str_sub(location, 1, 3)) 

```

# Cleaning data II
- Filtering out irrelevant data records

```{r}
# readr::guess_encoding("data/tptheft.csv")

url <- "https://data.taipei/api/getDatasetInfo/downloadResource?id=68785231-d6c5-47a1-b001-77eec70bec02&rid=93d9bc2d-af08-4db7-a56b-9f0a49226fa3"
df <- read_csv(url, locale = locale(encoding = "Big5")) %>% 
  select(date = "發生(現)日期", time = "發生時段", location = "發生(現)地點") %>%
  mutate(county = str_sub(location, 1, 3), 
         area = str_sub(location, 4, 5),
         year = date%/%10000,
         month = date%%10000%/%100,
         time = str_sub(time, 1, 2))%>% 
  filter(county == "臺北市")
```


# Long to wide form
- `count()` two variables
- `spread()` spread one variable as columns to wide form

```{r}
# count() then spread() df -> df.wide
df_wide <- df %>%
  count(area, time) %>% 
  spread(key = time, value = n, fill = 0) 

```


# Setting time as row.name for mosaicplot
- Setting rowname by `row.names()`
- Assigning NULL to variable in data.frame to delete it
```{r}
df_wide <- data.frame(df_wide)
row.names(df_wide) <- df_wide$area
df_wide$area <- NULL
```

```{r warning=FALSE}

# Specify fonts for Chinese
# par(family=('STKaiti')) 
par(family=('Heiti TC Light')) # for mac

# Specify colors
colors <- c('#D0104C', '#DB4D6D', '#E83015',  '#F75C2F',
            '#E79460', '#E98B2A', '#9B6E23', '#F7C242',
            '#BEC23F', '#90B44B', '#66BAB7', '#1E88A8')

# mosaicplot()
mosaicplot(df_wide, color=colors, border=0, off = 3,
           main="Theft rate of Taipei city (region by hour)")

```

#產假支薪

```{r}
library(tidyverse)
options(stringsAsFactors = F)
options(scipen = 999)
library(readxl)
read_excel("data/WORLD-MACHE_Gender_6.8.15.xls", "Sheet1", col_names=T) %>% 
    select(iso3, 6:24) %>%
    filter(matleave_13 == 5, matleave_95 != 5) %>% 
    gather("year", "degree", 2:20) %>% 
    #spread(year, degree, fill = 0) %>% View
    replace_na((list(degree = 0))) %>%
    mutate(year2 = as.POSIXct(strptime(year, "matleave_%y"))) %>% 
    mutate(year3 = lubridate::year(year2)) %>% 
    ggplot() +
    aes(year3, degree) +
    geom_col(color = "royalblue", fill = "royalblue") +
    facet_wrap(~ iso3) +
    theme_void()
    

```